---
title: "BDA - Project Work"
author: "Jacopo Losi, Nicola Saljoughi"
output:
  pdf_document:
    toc: yes
    toc_depth: 3
  word_document:
    toc: yes
    toc_depth: '1'
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '1'
---
```{r setup, include=FALSE}
# This chunk just sets echo = TRUE as default (i.e. print all code)
knitr::opts_chunk$set(echo = TRUE, tidy = FALSE)

library(magrittr)
#library(XLConnect)
library(rstan)
library(nnet)
library(epitools)
library(aaltobda)
library(tinytex)
library(MASS)
library(mvtnorm)
library(ggplot2)
library(rstan)
library(devtools)
library(brms)
library(loo)
library(KernSmooth)

data <- read.csv("./suicide attempt data_2.csv", stringsAsFactors=FALSE)

```

```{r}

mydata <- data
mydata$Season <- data$Month
mydata$Month = NULL

indexSpring <- which(data$Month >= 3 & data$Month <= 5)
indexSummer <- which(data$Month >= 6 & data$Month <= 8)
indexAutumn <- which(data$Month >= 9 & data$Month <= 11)
indexWinter <- which(data$Month == 12 | data$Month <= 2)

mydata$Season[indexSpring] <- '1'  # 1 --> Spring
mydata$Season[indexSummer] <- '2'  # 2 --> Summer
mydata$Season[indexAutumn] <- '3'  # 3 --> Autumn
mydata$Season[indexWinter] <- '4'  # 4 --> Winter

indexAgeOne   <- which(data$Age <= 34) 
indexAgeTwo   <- which(data$Age >= 35 & data$Age <= 49)
indexAgeThree <- which(data$Age >= 50 & data$Age <= 64)
indexAgeFour  <- which(data$Age >= 65)

mydata$Age[indexAgeOne]   <- '1'   # 1 --> <34
mydata$Age[indexAgeTwo]   <- '2'   # 2 --> 35-49
mydata$Age[indexAgeThree] <- '3'   # 3 --> 50-64
mydata$Age[indexAgeFour]  <- '4'   # 4 --> >65

indexFarm   <- which(data$Occupation == 'farming')
indexNoFarm <- which(data$Occupation != 'farming')

mydata$Occupation[indexFarm]   <- 1    # 1 --> framing
mydata$Occupation[indexNoFarm] <- 2    # 2 --> non farming

indexOthers <- which(data$method != 'Pesticide' &
                     data$method != 'Other poison' &
                     data$method != 'hanging')

mydata$method[indexOthers] <- 'All others'

```


```{r}
```

# Introduction


## Objective

The objective of the study is to estimate the incidence of seriours suicide attempts (SSAs), defined as suicide attempts resulting in either death or hospitalization, and to analyse the factors associated with fatality among the attempters. 

## Data 
The data set is constituted by 2571 observations of 11 variables:
\begin{itemize}
  \item \texttt{Person\_ID}: ID number, $1,...,2571$
  \item \texttt{Hospitalised}: \textit{yes} or \textit{no}
  \item \texttt{Died}: \textit{yes} or \textit{no}
  \item \texttt{Urban}: \textit{yes}, \textit{no} or \textit{unknown}
  \item \texttt{Year}: $2009$, $2010$ or $2011$
  \item \texttt{Month}: $1,...,12$
  \item \texttt{Sex}: \textit{female} or \textit{male}
  \item \texttt{Age}: years
  \item \texttt{Education}: \textit{iliterate}, \textit{primary}, \textit{Secondary}, \textit{Tertiary} or \textit{unknown}
  \item \texttt{Occupation}: one of ten categories
  \item \texttt{method}: one of nine methods
\end{itemize}

## Source

Sun J, Guo X, Zhang J, Wang M, Jia C, Xu A (2015) "Incidence and fatality of serious suicide attempts in a predominantly rural population in Shandong, China: a public health surveillance study," BMJ Open 5(2): e006762. https://doi.org/10.1136/bmjopen-2014-006762

Data downloaded via Dryad Digital Repository. https://doi.org/10.5061/dryad.r0v35


# Analysis

```{r}

rural_men   <- subset(data, data$Sex=="male" & data$Urban=="no")
rural_women <- subset(data, data$Sex=="female" & data$Urban=="no")
urban_men   <- subset(data, data$Sex=="male" & data$Urban=="yes")
urban_women <- subset(data, data$Sex=="female" & data$Urban=="yes")

```
```{r}

str(data)

```


```{r}

## Create Stan data
dat <- list(N        = nrow(data),
            p        = 10,
            died     = data$Died,
            urban    = data$Urban,
            year     = data$Year,
            month    = data$Season,
            sex      = data$Sex,
            age      = data$Age,
            edu      = data$Education,
            job      = data$Occupation,
            method   = data$method)

## Load Stan file
fileName <- "./logistic_regression_model.stan"
stan_code <- readChar(fileName, file.info(fileName)$size)
cat(stan_code)

```


\begin{tabular}{p{13mm} p{10mm} p{25mm} p{20mm} p{25mm} p{35mm} p{20mm} }
\hline
  & \textbf{All SAAs}  &  \textbf{Hospitalised} \newline \textbf{and survived}  &  \textbf{Hospitalised} \newline \textbf{but died}  &  \textbf{Total SSA} \newline \textbf{hospitalisations} & \textbf{SSA deaths without} \newline \textbf{hospitalisation} & \textbf{Total SSA} \newline \textbf{deaths} \\
  \hline
  Urban      &      &      &     &      &      &      \\
  \;\;Female & 149  & 99   & 18  & 117  & 32   & 50   \\
  \;\;Male   & 128  & 65   & 17  & 82   & 46   & 63   \\
  \;\;Both   & 277  & 164  & 35  & 199  & 78   & 113  \\
  Rural      &      &      &     &      &      &      \\
  \;\;Female & 1134 & 598  & 100 & 698  & 436  & 536  \\
  \;\;Male   & 1079 & 474  & 103 & 577  & 502  & 605  \\
  \;\;Both   & 2213 & 1072 & 203 & 1275 & 938  & 1141 \\
  Total      &      &      &     &      &      &      \\
  \;\;Female & 1328 & 741  & 118 & 859  & 469  & 587  \\
  \;\;Male   & 1243 & 574  & 120 & 694  & 549  & 669  \\
  \;\;Both   & 2571 & 1315 & 238 & 1553 & 1018 & 1256 \\
  \hline
  
\end{tabular}


# Conclusions
